---
title: "lab 2"
author: "Nate"
date:
output:
  html_document:
    theme: flatly
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, out.width = "75%", fig.align = 'center')
```

# Question:1

### Step 1
##### Covid Url (Read In The Url)
```{r, message=FALSE,warning=FALSE}
library(tidyverse)
covid=read_csv("https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-counties.csv")
```

### Step 2
```{r, message=FALSE,warning=FALSE}
q1 = covid %>% 
  filter(state == "California")%>%
  group_by(county) %>% 
  mutate(newCases = c(cases[1], diff(cases))) %>% 
  ungroup() %>% 
  filter(date >= max(date)-13)
```

Making A Subset That Filters The Data To California And Added A New Column (Mutate) With The Daily New Cases Using Either diff or lag.

### Step 3

```{r}
most_all_time = q1 %>% 
  filter(date == max(date)) %>% 
  slice_max(cases, n = 5) %>% 
  select(county, cases)

knitr::kable(most_all_time, caption = "Most Cumulative Cases / Most Cases Of All Time", col.names = c("Counties", "Max Cases To Date"))
```

Five Counties with the Most Cumulative Case

```{r}
most_new_data_from_yesterday = q1 %>% 
  filter(date == max(date)) %>% 
  slice_max(newCases, n = 5)

knitr::kable(most_all_time, caption = "Five Counties With The Most New Data (From Yesterday)", col.names = c("Counties", "New Cases"))
```

### Step 4

Please download the data and store it in the data directory of your project. = Done!
  
### Step 5

Loaded the population data with the “dataset importer” (it was found in the file in my data directory via the file explorer –> click on it –> “Import Dataset”. I copied the code preview (ignored the View(...)) and inserted it in this Rmarkdown. This will allow the data to be referenced every time the file is run! = Done!

### Step 6

```{r}
library(readxl)

PopulationEstimates=read_excel("../data/PopulationEstimates.xls", 
  skip = 2) %>% 
  select(fips = FIPStxt, pop2019 = "POP_ESTIMATE_2019", Area_Name)
```

### Step 7
```{r}
j1 = left_join(covid, PopulationEstimates, by = "fips")
```

Join the population data to the California COVID data.

### Step 9
##### Number Of Cases In The Last 14 Days Per 100,000 

```{r, message=FALSE,warning=FALSE}
last14 = j1 %>% 
  filter(state ==c("California")) %>% 
  group_by(county, date) %>% 
  summarize(totalcases = sum(cases, na.rm = TRUE),
            pop2019    = sum(pop2019, na.rm = TRUE)) %>%
  ungroup() %>% 
  filter(date >= max(date) - 13) %>% 
  group_by(county, pop2019) %>% 
  mutate(newcases = totalcases-lag(totalcases)) %>% 
  summarize(totNewCases = sum(newcases, na.rm = TRUE)) %>% 
  mutate(per100 = totNewCases / (pop2019/100000)) %>% 
  filter(per100 <= 100)
```

# Question 2

In this question, we looked at the story of 4 states and the impact scale it can have on data interpretation. The states include: New York, California, Louisiana, and Florida. My task was to make a faceted bar plot showing the number of daily, new cases at the state level.

### Step 1

```{r, message=FALSE,warning=FALSE}

fourStates=covid %>% 
  group_by(state,date) %>% 
  summarize(cases=sum(cases)) %>%
  ungroup() %>% 
  filter(state %in% c("California", "New York", "Florida", "Louisiana")) %>%
  group_by(state) %>% 
  mutate(newCases=cases-lag(cases)) %>%
  mutate(roll7=zoo::rollmean(newCases,7,fill=NA,
                      align="right"))

```

### Step 2

```{r, message=FALSE,warning=FALSE}
ggplot(data = fourStates, aes(x = date)) + 
  geom_col(aes(y = newCases), col = "pink") + 
  geom_line(aes(y= roll7), col = "darkred")+
  facet_wrap(~state) +
  labs(title = "Daily New Cases",  
       x = "Date",  
       y = "New Cases",  
       caption = "Based On NY Times Covid Data",  
       subtitle = 'Data Source: NY Times') +
       theme_bw()
```

```{r, message=FALSE,warning=FALSE}
new = fourStates %>% left_join(PopulationEstimates, by = c("state" = "Area_Name") ) %>% 
  mutate(perCapNew = newCases / pop2019, 
         perCap7 = zoo::rollmean(perCapNew, 7, fill = NA, align = "center" ))


ggplot(data = new, aes(x = date)) + 
  geom_col(aes(y = perCapNew), col = "pink") + 
  geom_line(aes(y= perCap7), col = "darkred")+
  facet_wrap(~state) +
  labs(title = "Daily New Cases",  
       x = "Date",  
       y = "New Cases",  
       caption = "Based On NY Times Covid Data",  
       subtitle = 'Data Source: NY Times') +
       theme_bw()
```

